---
title: "EDA of DC Zillow Listings Vs AirBnB"
author: "Bryan Johns"
date: "2025-02"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(devtools)
library(tidyverse)
library(scales)
library(sf)
library(ggmap)
library(ggspatial)
library(rnaturalearth)
library(rnaturalearthdata)
load_all("../..") # packages that clean and prep the data for analysis / visualization
```

# DC AirBnB Vs Zillow listings

This is a brief exploratory data analysis of AirBnB listings vs Zillow listings for Washington, DC, using web-scraped data from roughly June of 2024. There are no real interesting observations. I was trying out R (this is my first ever R Markdown) as part of the requirements for the final project of [cs50's Introduction to Programming with R](https://cs50.harvard.edu/r/2024/).

I'm playing with R visualizations.

## Data Preparation and Cleaning

Loading the Zillow data:

```{r, echo=FALSE}
re_df <- read_real_estate_json()
re_df <- as_tibble(re_df)
head(re_df)
```
Loading the AirBnB data, which drops a few errors ($7000 hostel beds that should be $70):

```{r, echo=FALSE}
adf <- read_airbnb_csv()
head(adf)
```

Assign neighborhoods to the Zillow data. This drops 6 rows lacking lat/long coordinates.

```{r, echo=FALSE}
zdf <- assign_neighborhood(re_df)
zdf <- as_tibble(zdf)
head(zdf)
```
Verify that neighborhoods have been assigned correctly. The black plots are in DC, Red in Virginia and Maryland.

```{r, echo=FALSE}
ggplot(zdf, aes(x = longitude, y = latitude, color = neighborhood == "Unknown")) +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  labs(title = "Spatial Distribution of Zillow Listings",
       subtitle = "Checking that listings are assigned to DC neighborhoods",
       color = "Unknown Neighborhood") +
  theme_minimal()
```

The red plots are "Unknown" in the upcoming Neighborhood counts:

```{r, echo=FALSE}
zdf %>%
  count(shortNeighborhood, sort = TRUE)
```

## Map Zillow and AirBnB properties by neighborhood

Drop "Unknown" neighborhoods and see if neighborhoods coded correctly:

```{r, echo=FALSE}
zdf <- zdf[zdf$neighborhood != "Unknown", ]
ggplot(zdf, aes(x = longitude, y = latitude, color = neighborhood)) +
  geom_point(alpha = 0.7) +
  labs(title = "Spatial Distribution of Zillow Listings",
       subtitle = "Checking that listings are assigned to DC neighborhoods",
       color = "Neighborhood") +
  theme_minimal() +
  theme(legend.position = "none") # neighborhood names squish out the plot
```

And the AirBnB's:

```{r, echo=FALSE}
ggplot(adf, aes(x = longitude, y = latitude, color = neighborhood)) +
  geom_point(alpha = 0.7) +
  labs(title = "Spatial Distribution of AirBnB Listings",
       # subtitle = "Checking that listings are assigned to DC neighborhoods",
       color = "Neighborhood") +
  theme_minimal() +
  theme(legend.position = "none") # neighborhood names squish out the plot
```

## Average Price

```{r, echo=FALSE}
# calculate average price by neighborhood for zdf...
zdf_avg_price <- zdf %>%
  group_by(shortNeighborhood) %>%
  summarize(avg_price_zillow = mean(price, na.rm = TRUE))

# ...and adf
adf_avg_price <- adf %>%
  group_by(shortNeighborhood) %>%
  summarize(avg_price_airbnb = mean(price, na.rm = TRUE))

# join as a df
combined_avg_price <- full_join(zdf_avg_price, adf_avg_price, by = "shortNeighborhood")

# get total average price, add to df
total_avg_price_zillow <- mean(zdf$price, na.rm = TRUE)
total_avg_price_airbnb <- mean(adf$price, na.rm = TRUE)
combined_avg_price <- combined_avg_price %>%
  add_row(shortNeighborhood = "Total", avg_price_zillow = total_avg_price_zillow, avg_price_airbnb = total_avg_price_airbnb)

# calculate percent change from total average price for each neighborhood
combined_avg_price <- combined_avg_price %>%
  mutate(
    change_zillow = (avg_price_zillow - total_avg_price_zillow) / total_avg_price_zillow * 100,
    change_airbnb = (avg_price_airbnb - total_avg_price_airbnb) / total_avg_price_airbnb * 100
  )

# format
combined_avg_price_formatted <- combined_avg_price %>%
  mutate(
    avg_price_zillow = dollar(avg_price_zillow, accuracy = 0.01),
    avg_price_airbnb = dollar(avg_price_airbnb, accuracy = 0.01),
    change_zillow = percent(change_zillow / 100, accuracy = 0.01),
    change_airbnb = percent(change_airbnb / 100, accuracy = 0.01)
  )

print(combined_avg_price_formatted)
```

Create a scatterplot comparing percent change in Zillow versus percent change in Airbnb:

```{r, echo=FALSE}
ggplot(combined_avg_price, aes(x = avg_price_zillow, y = avg_price_airbnb)) +
  geom_point() +
  # geom_text(aes(label = shortNeighborhood), vjust = -0.5, hjust = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Zillow vs Airbnb Prices by Neighborhood",
       x = "Zillow Prices",
       y = "Airbnb Prices") +
  theme_minimal()
```


It doesn't look like much of a correlation. Let's prove it. Test for fit:

```{r, echo=FALSE}
# Fit a linear model
fit <- lm(avg_price_airbnb ~ avg_price_zillow, data = combined_avg_price)

# Print the summary of the linear model
summary(fit)
```

Conclusion:

AirBnB prices have little to do with Zillow prices (on the neighborhood level) - perhaps because the neighborhoods are not subdivided enough, perhaps because vacationers are looking for different things than home buyers?

## Price Histograms

```{r, echo=FALSE}
adf_filtered <- adf %>% filter(adf$price < 1000)

ggplot() +
  geom_histogram(data = adf_filtered, aes(x = price, fill = "Airbnb"), binwidth = 50, alpha = 0.5, position = "identity") +
  scale_fill_manual(values = c("Airbnb" = "red")) +
  labs(title = "Histogram of Airbnb Prices", x = "Price", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r, echo=FALSE}
zdf_filtered <- zdf %>% filter(zdf$price < 2000000)

ggplot() +
  geom_histogram(data = zdf_filtered, aes(x = price, fill = "Zillow"), binwidth = 50000, alpha = 0.5, position = "identity") +
  scale_fill_manual(values = c("Zillow" = "blue")) +
  labs(title = "Histogram of Zillow Prices", x = "Price", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")
```


```{r, echo=FALSE}
ggplot() +
  # geom_boxplot(data = adf, aes(x = shortNeighborhood, y = price, fill = "Airbnb"), alpha = 0.5) +
  geom_boxplot(data = zdf, aes(x = shortNeighborhood, y = price, fill = "Zillow"), alpha = 0.5) +
  scale_fill_manual(values = "blue") +
  labs(title = "Airbnb vs Zillow Prices by Neighborhood", x = "Neighborhood", y = "Price") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()
  # theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) # +
  # coord_flip()
```

```{r, echo=FALSE}
ggplot() +
  geom_boxplot(data = adf, aes(x = shortNeighborhood, y = price, fill = "Airbnb"), alpha = 0.5) +
  # geom_boxplot(data = zdf, aes(x = shortNeighborhood, y = price, fill = "Zillow"), alpha = 0.5) +
  scale_fill_manual(values = "red") +
  labs(title = "Airbnb vs Zillow Prices by Neighborhood", x = "Neighborhood", y = "Price") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()
```

## Median Price

```{r, echo=FALSE}
airbnb_summary <- adf %>%
  group_by(shortNeighborhood) %>%
  summarise(median_airbnb_price = median(price, na.rm = TRUE))
airbnb_summary
```

```{r, echo=FALSE}
zillow_summary <- zdf %>%
  group_by(shortNeighborhood) %>%
  summarise(median_zillow_price = median(price, na.rm = TRUE))
zillow_summary
```

Heatmap of Price by Neighborhood:

```{r, echo=FALSE}
# normalize the prices for the heatmap
airbnb_summary <- airbnb_summary %>%
  mutate(median_a_scaled = scale(median_airbnb_price))

zillow_summary <- zillow_summary %>%
  mutate(median_z_scaled = scale(median_zillow_price))

# combine and pivot longer
heatmap_df <- full_join(
  airbnb_summary %>% select(shortNeighborhood, median_a_scaled),
  zillow_summary %>% select(shortNeighborhood, median_z_scaled),
  by = "shortNeighborhood"
) %>%
  pivot_longer(cols = c(median_a_scaled, median_z_scaled),
               names_to = "Type", values_to = "Price")

# create heatmap
ggplot(heatmap_df, aes(x = shortNeighborhood, y = Type, fill = Price)) +
  geom_tile() +
  scale_fill_gradient(low = "yellow", high = "red") +
  theme_minimal() +
  labs(title = "Heatmap of Scaled Median Airbnb vs Zillow Prices", x = "Neighborhood", y = "Type") +
  coord_flip() # +
  # theme(axis.text.x = element_text(angle = 90, hjust = 1))  # Rotate x-axis labels for legibility
```

```{r, echo=FALSE}
roi_df <- full_join(airbnb_summary, zillow_summary, by = "shortNeighborhood") %>%
  mutate(percent = (median_airbnb_price / median_zillow_price) * 100)

ggplot(roi_df, aes(x = reorder(shortNeighborhood, percent), y = percent)) +
  geom_col(fill = "blue") +
  coord_flip() +
  labs(title = "Airbnb Price as % of Zillow Price", x = "Neighborhood", y = "Percentage (%)") +
  theme_minimal()

```

## Listings per Neighborhood

```{r, echo=FALSE}
airbnb_counts <- as.data.frame(table(adf$shortNeighborhood))
colnames(airbnb_counts) <- c("shortNeighborhood", "airbnb_listings")

zillow_counts <- as.data.frame(table(zdf$shortNeighborhood))
colnames(zillow_counts) <- c("shortNeighborhood", "zillow_listings")

# Merge counts on neighborhood
listings_count <- full_join(airbnb_counts, zillow_counts, by = "shortNeighborhood") %>%
  replace_na(list(airbnb_listings = 0, zillow_listings = 0)) %>%
  pivot_longer(cols = c(airbnb_listings, zillow_listings), names_to = "Listing Type", values_to = "Count")

# Plot
ggplot(listings_count, aes(x = reorder(shortNeighborhood, Count), y = Count, fill = `Listing Type`)) +
  geom_col(position = "dodge") +
  # coord_flip() +
  scale_fill_manual(values = c("airbnb_listings" = "red", "zillow_listings" = "blue")) +
  labs(title = "Number of Listings: Airbnb vs Zillow", x = "Neighborhood", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r, echo=FALSE}
airbnb_types <- adf %>%
  group_by(shortNeighborhood, room_type) %>%
  summarise(count = n(), .groups = "drop")

ggplot(airbnb_types, aes(x = shortNeighborhood, y = count, fill = room_type)) +
  geom_bar(stat = "identity") +
  # coord_flip() +
  scale_fill_viridis_d() +
  labs(title = "Airbnb Property Types by Neighborhood", x = "Neighborhood", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r, echo=FALSE}

# Filter adf for prices less than 500
adf_filtered <- adf %>% filter(price < 500)
ggplot(adf_filtered, aes(x = shortNeighborhood, y = number_of_reviews, fill = price)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(title = "Heatmap of Reviews vs AirBnB Prices", x = "Neighborhood", y = "Number of Reviews") +
  coord_flip()
```

## Hypothetical ROI

Let's estimate the Return on Investment, if hypothetically buying a house to rent out as an AirBnB, using reviews per month. Note that this is extremely crude and assumption laden. Neighborhoods are large blocks and contain diversity within them, no granular detail included on type of rental (whole property? shared room?), and an off-the-cuff two nights per review (a possible way to refine this estimate would be based on future availability, itself an assumption-laden approach).

```{r, echo=FALSE}
roi_table <- adf %>%
  group_by(shortNeighborhood) %>%
  summarise(avg_airbnb_price = mean(price, na.rm = TRUE),
            avg_reviews = mean(reviews_per_month, na.rm = TRUE)) %>%
  full_join(zillow_summary, by = "shortNeighborhood") %>%
  mutate(estimated_revenue = avg_airbnb_price * avg_reviews * 12 * 2, # 12 months, 2 nights per review
         roi = estimated_revenue / median_zillow_price) %>%
  select(shortNeighborhood, roi, everything()) %>% # reorder columns
  arrange(desc(roi))

# format roi as a percent
roi_table_percent <- roi_table %>%
  mutate(
    roi = percent(roi),
    avg_airbnb_price = dollar(avg_airbnb_price, accuracy = 0.01),
    median_zillow_price = dollar(median_zillow_price, accuracy = 0.01),
    estimated_revenue = dollar(estimated_revenue, accuracy = 0.01)
  )

print(roi_table_percent)

```

ROI per neighborhood distribution:

```{r, echo=FALSE}
summary(roi_table$roi)
```

## Some Maps

```{r, echo=FALSE}
# load the geojson, note the British spelling of neighbourhood
neighborhoods <- st_read(system.file("extdata", "neighbourhoods_cleaned.geojson", package = "DC_property_data"), quiet = TRUE) 

# aggregate review counts by neighborhood
reviews_by_neighborhood <- adf %>%
  group_by(neighbourhood = neighborhood) %>%
  summarize(total_reviews = sum(number_of_reviews, na.rm = TRUE))

# merge review data with spatial neighborhoods
choropleth_data <- neighborhoods %>%
  left_join(reviews_by_neighborhood, by = "neighbourhood")

# create the choropleth map
ggplot() +
  geom_sf(data = choropleth_data, aes(fill = total_reviews), color = "black") +
  scale_fill_gradient(low = "blue", high = "red", na.value = "grey50") +
  theme_void() +
  labs(title = "Choropleth of Airbnb Reviews by Neighborhood", fill = "Total Reviews") +
  theme(legend.position = "right")
```

```{r, echo=FALSE}
# load the geojson, note the British spelling of 'neighbourhood'
neighborhoods <- st_read(system.file("extdata", "neighbourhoods_cleaned.geojson", package = "DC_property_data"), quiet = TRUE) 

# aggregate median price by neighborhood
price_by_neighborhood <- adf %>%
  group_by(neighbourhood = neighborhood) %>%  # Match column name in GeoJSON
  summarize(median_price = median(price, na.rm = TRUE))

# merge price data with spatial neighborhoods
choropleth_data <- neighborhoods %>%
  left_join(price_by_neighborhood, by = "neighbourhood")

# create the choropleth map
ggplot() +
  geom_sf(data = choropleth_data, aes(fill = median_price), color = "black") +
  scale_fill_gradient(low = "blue", high = "red", na.value = "grey50") +
  theme_void() +
  labs(title = "Choropleth of Airbnb Median Price by Neighborhood", fill = "Median Price") +
  theme(legend.position = "right")
```



```{r, echo=FALSE}
# convert adf to sf object
adf_sf <- st_as_sf(adf, coords = c("longitude", "latitude"), crs = 4326)

# get basemap from OpenStreetMap
ggplot() +
  annotation_map_tile(type = "osm", zoom = 12) +  # getch OSM basemap
  geom_sf(data = adf_sf, aes(size = number_of_reviews, color = price), alpha = 0.6) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Bubble Map: Popularity vs AirBnB Prices", subtitle = "Based on number of reviews", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r, echo=FALSE}
# aggregate data by neighborhood
neighborhood_summary <- adf %>%
  group_by(neighborhood) %>%
  summarize(
    total_reviews = sum(number_of_reviews, na.rm = TRUE),
    median_price = median(price, na.rm = TRUE),
    longitude = mean(longitude, na.rm = TRUE), 
    latitude = mean(latitude, na.rm = TRUE)
  ) %>%
  drop_na(longitude, latitude)  # remove any NA coordinates

# convert to sf object
neighborhood_sf <- st_as_sf(neighborhood_summary, coords = c("longitude", "latitude"), crs = 4326)

# plot with OpenStreetMap basemap
ggplot() +
  annotation_map_tile(type = "osm", zoom = 12) +  # fetch OSM basemap
  geom_sf(data = neighborhood_sf, aes(size = total_reviews, color = median_price), alpha = 0.7) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Bubble Map: Popularity vs Airbnb Prices by Neighborhood",
       subtitle = "Bubble size = Total Reviews, Color = Median Price",
       x = "Longitude", y = "Latitude") +
  theme_minimal()
```

